Iterative construction of replicated designs based on Sobol' sequences

### 2016-06-14 crass/crassnew.tex
Big changes in definition of oplus and algo additive approach

@@

$s$-dimensional Sobol' sequences are digital sequences in base $b=2$ that can be computed using the \emph{generating matrices}, a set of $s$ full rank infinite dimensional upper triangular matrices over the Galois field $\mathbb{F}_2:=\{0,1\}$. We will denote by $\oplus$ the addition in $\mathbb{F}_2$. These generating matrices are recursively constructed given some primitive polynomials and initial directional numbers. In \cite{Kuo}, Kuo and Joe detail this construction and also suggest a particular choice for these matrices that optimize the 2 dimensional projection $t$-values.

@@

@@

At step $\ell$, $\ell \geq 0$, $B_\ell=\{\vx_i^{(\ell)}\}_{i=0}^{2^r-1}$ is generated as follows: for each $i=0,\dots,2^r-1$, point $\vx_i^{(\ell)} = (x_{i,1}^{(\ell)},\dots,x_{i,s}^{(\ell)})^\intercal$ is obtained by linearly transforming $\vx_i$,
\begin{equation}
\label{dig.shift}
(x_{i,j,1}^{(\ell)},\dots,x_{i,j,r}^{(\ell)})^\intercal = L (x_{i,j,1},\dots,x_{i,j,r})^\intercal\oplus \ve_j,\qquad j= 1,\dots,s\, ,
\end{equation}
where $ x_{i,j}^{(\ell)}=\sum_{k = 1}^r x_{i,j,k}^{(\ell)}2^{-k}$, $\ve_j=(e_{j,{0}},\dots,e_{j,{r-1}})^\intercal\in\mathbb{F}_2^r$ and $L$ is a full rank lower triangular matrix of size $r \times r$ over $\mathbb{F}_2$.
\bigskip

Likewise, ${B'}_\ell=\{{\vx'}_i^{(\ell)}\}_{i=0}^{2^r-1}$ is obtained as follows: for each $i=0,\dots,2^r-1$, point ${\vx'}_i^{(\ell)} = ({x'}_{i,1}^{(\ell)},\dots,{x'}_{i,s}^{(\ell)})^\intercal$ is obtained by linearly transforming ${\vx'}_i^{(0)}$,
\begin{equation}
\label{ibinom.scrambling}
({x'}_{i,j,1}^{(\ell)},\dots,{x'}_{i,j,r}^{(\ell)})^\intercal = {L'} ({x'}_{i,j,1},\dots,{x'}_{i,j,r})^\intercal\oplus {\ve'}_j,\qquad j= 1,\dots,s\, ,
\end{equation}
where $ {x'}_{i,j}^{(\ell)}=\sum_{k = 1}^r {x'}_{i,j,k}^{(\ell)}2^{-k}$, ${\ve'}_j=(e_{j,{0}},\dots,e_{j,{r-1}})^\intercal\in\mathbb{F}_2^r$ and ${L'}$ is a full rank lower triangular matrix of size $r \times r$ over $\mathbb{F}_2$.
\bigskip

Left multiplications by matrices $L$ and $L'$ in \eqref{dig.shift} and \eqref{ibinom.scrambling}  are called linear scramblings \cite{Matousek} while additions by vectors $\ve_j$ and ${\ve'}_j$ are referred as digital shifts. There are exactly $2^r$ different choices of $\ve_j$ and ${\ve'}_j$ and $2^{r(r-1)/2}$ choices of $L$ and $L'$. To guarantee that new points added will fill unexplored regions of the input space, the vectors $\ve_j$ are selected as follows:
\begin{itemize}
\item[i)] at iteration $\ell=0$, vectors $\ve_j$ are all equal to the zero of $\mathbb{F}_2^r$,
\item[ii)] for $\ell \geq1$, we use the following result:
$(B_0,\oplus)$ is a subgroup of $\left(\mathbb{F}_2^{r}\right)^{s}$. From Lagrange's Theorem, we can construct cosets $B_\ell$ as $B_0 \oplus \ve$, $\ve=(\ve_1,\dots,\ve_s) \in \left(\mathbb{F}_2^{r}\right)^{s}$
%, which form a partition of $\left(\mathbb{F}_2^{r}\right)^{s}$
. At iteration $\ell \geq1$, we select $\ve$ such that $B_0 \oplus \ve$ is a new coset different from $\{B_0,\dots,B_{\ell-1}\}$, and it will be enough to choose $\ve\notin\mathcal{P}_{\ell-1}$. By identification we have $B_{\ell}=B_0 \oplus \ve$ ensuring that the new points added are all different.

%\item[i)] for the first dimension, the same vector $\ve_1=(0,\dots,0)$ (resp. ${\ve}'_1=(0,\dots,0)$) is selected at each iteration $\ell$,
%\item[ii)] for the other dimensions, vectors $\ve_2,\dots,\ve_s$ (resp. ${\ve}'_2,\dots,{\ve}'_s$) are selected to guarantee that the set $\{\ve_2,\dots,\ve_s\}$ is different at each step $\ell$ and different from $\{\ve_1,\dots,\ve_1\}$. There exists exactly $2^{r (s-1)}$ combinations of this set. At each step $\ell$, one combination is sampled without replacement. 
\end{itemize}
The same reasoning to select vectors $\ve'_j$ is made for $({B'}_0,\oplus)$. The maximum number of iterations $\ell$ that can be performed equals the number of cosets, $2^{s}$. If this maximum is reached, designs $\mathcal{P}_\ell$ and $\mathcal{P}'_\ell$ both correspond to a full grid design of $2^{rs}$ points.
%We have the following result:
%\item[Step 4.] The two blocks $B_l$ and ${B'}_l$ are then both randomized using the same Owen's scrambling column-wise.

>>>>

Each

@@

Sobol' sequences in dimension $s$ are digital sequences in base $b=2$ that can be computed using the \emph{generating matrices}, a set of $s$ full rank infinite dimensional upper triangular matrices over the Galois field $\mathbb{F}_2:=\{0,1\}$.
%We will denote by $\oplus$ the addition in $\mathbb{F}_2$.
These generating matrices are recursively constructed given some primitive polynomials and initial directional numbers. In \cite{Kuo}, Kuo and Joe detail this construction and also suggest a particular choice for these matrices that optimize the 2 dimensional projection $t$-values.

@@

Sobol' sequences also have good group structure properties. For any $m\geq 1$, the first $2^m$ points of the sequence form an Abelian group under the following operation $\oplus$,
\begin{gather*}
\vx = \left(\sum_{k=1}^m x_{1,k}2^{-k},\dots,\sum_{k=1}^m x_{s,k}2^{-k}\right)^\intercal\; , \qquad \vz = \left(\sum_{k=1}^m z_{1,k}2^{-k},\dots,\sum_{k=1}^m z_{s,k}2^{-k}\right)^\intercal, \\
\vx\oplus\vz := \left(\sum_{k=1}^m (x_{1,k}+z_{1,k}\mod 2)2^{-k},\dots,\sum_{k=1}^m (x_{s,k}+z_{s,k}\mod 2)2^{-k}\right)^\intercal.
\end{gather*}

@@

%At step $\ell$, $\ell \geq 0$, $B_\ell=\{\vx_i^{(\ell)}\}_{i=0}^{2^r-1}$ is generated as follows: for each $i=0,\dots,2^r-1$, point $\vx_i^{(\ell)} = (x_{i,1}^{(\ell)},\dots,x_{i,s}^{(\ell)})^\intercal$ is obtained by linearly transforming $\vx_i$,
%\begin{equation}
%\label{dig.shift}
%(x_{i,j,1}^{(\ell)},\dots,x_{i,j,r}^{(\ell)})^\intercal = L (x_{i,j,1},\dots,x_{i,j,r})^\intercal\oplus \ve_j,\qquad j= 1,\dots,s\, ,
%\end{equation}
%where $ x_{i,j}^{(\ell)}=\sum_{k = 1}^r x_{i,j,k}^{(\ell)}2^{-k}$, $\ve_j=(e_{j,{0}},\dots,e_{j,{r-1}})^\intercal\in\mathbb{F}_2^r$ and $L$ is a full rank lower triangular matrix of size $r \times r$ over $\mathbb{F}_2$.
%\bigskip
%
%Likewise, ${B'}_\ell=\{{\vx'}_i^{(\ell)}\}_{i=0}^{2^r-1}$ is obtained as follows: for each $i=0,\dots,2^r-1$, point ${\vx'}_i^{(\ell)} = ({x'}_{i,1}^{(\ell)},\dots,{x'}_{i,s}^{(\ell)})^\intercal$ is obtained by linearly transforming ${\vx'}_i^{(0)}$,
%\begin{equation}
%\label{ibinom.scrambling}
%({x'}_{i,j,1}^{(\ell)},\dots,{x'}_{i,j,r}^{(\ell)})^\intercal = {L'} ({x'}_{i,j,1},\dots,{x'}_{i,j,r})^\intercal\oplus {\ve'}_j,\qquad j= 1,\dots,s\, ,
%\end{equation}
%where $ {x'}_{i,j}^{(\ell)}=\sum_{k = 1}^r {x'}_{i,j,k}^{(\ell)}2^{-k}$, ${\ve'}_j=(e_{j,{0}},\dots,e_{j,{r-1}})^\intercal\in\mathbb{F}_2^r$ and ${L'}$ is a full rank lower triangular matrix of size $r \times r$ over $\mathbb{F}_2$.
%\bigskip

Initial sets are $B_0=\{\vx_i\}_{i=0}^{2^r-1}$ and ${B'}_0=\{{\vx'}_i\}_{i=0}^{2^r-1}$. Then, one may take $B_\ell=\{\vx_i\oplus\ve_\ell\}_{i=0}^{2^r-1}$ and ${B'}_\ell=\{{\vx'}_i\oplus{\ve'}_\ell\}_{i=0}^{2^r-1}$ such that $\ve_\ell\notin\mathcal{P}_{\ell-1}$ and ${\ve'}_\ell\notin\mathcal{P'}_{\ell-1}$.  Indeed, $B_0$ is a subgroup of $\left(\mathbb{F}_2^{r}\right)^{s}$. Therefore, we can construct cosets $B_\ell$ as $B_0 \oplus \ve$, under the right $\ve \in \left(\mathbb{F}_2^{r}\right)^{s}$. After that, we may also apply linear scramblings \cite{Matousek} using lower triangular matrices $L$ and $L'$ ...... There are exactly $2^{rs}$ different choices of $\ve$ and ${\ve'}$, and $2^{r(r-1)/2}$ choices of $L$ and $L'$. The maximum number of iterations $\ell$ that can be performed equals the number of cosets, $2^{s}$. If this maximum is reached, designs $\mathcal{P}_\ell$ and $\mathcal{P}'_\ell$ both correspond to a full grid design of $2^{rs}$ points.

%Left multiplications by matrices $L$ and $L'$ in \eqref{dig.shift} and \eqref{ibinom.scrambling}  are called linear scramblings \cite{Matousek} while additions by vectors $\ve_j$ and ${\ve'}_j$ are referred as digital shifts. There are exactly $2^r$ different choices of $\ve_j$ and ${\ve'}_j$ and $2^{r(r-1)/2}$ choices of $L$ and $L'$. To guarantee that new points added will fill unexplored regions of the input space, the vectors $\ve_j$ are selected as follows:
%\begin{itemize}
%\item[i)] at iteration $\ell=0$, vectors $\ve_j$ are all equal to the zero of $\mathbb{F}_2^r$,
%\item[ii)] for $\ell \geq1$, we use the following result:
%$(B_0,\oplus)$ is a subgroup of $\left(\mathbb{F}_2^{r}\right)^{s}$. From Lagrange's Theorem, we can construct cosets $B_\ell$ as $B_0 \oplus \ve$, $\ve=(\ve_1,\dots,\ve_s) \in \left(\mathbb{F}_2^{r}\right)^{s}$
%%, which form a partition of $\left(\mathbb{F}_2^{r}\right)^{s}$
%. At iteration $\ell \geq1$, we select $\ve$ such that $B_{\ell}=B_0 \oplus \ve$ is a new coset different from $\{B_0,\dots,B_{\ell-1}\}$. It will be enough to choose $\ve$ such that $\ve\notin\mathcal{P}_{\ell-1}$. %By identification we have $B_{\ell}=B_0 \oplus \ve$ ensuring that the new points added are all different.

%\item[i)] for the first dimension, the same vector $\ve_1=(0,\dots,0)$ (resp. ${\ve}'_1=(0,\dots,0)$) is selected at each iteration $\ell$,
%\item[ii)] for the other dimensions, vectors $\ve_2,\dots,\ve_s$ (resp. ${\ve}'_2,\dots,{\ve}'_s$) are selected to guarantee that the set $\{\ve_2,\dots,\ve_s\}$ is different at each step $\ell$ and different from $\{\ve_1,\dots,\ve_1\}$. There exists exactly $2^{r (s-1)}$ combinations of this set. At each step $\ell$, one combination is sampled without replacement. 
%\end{itemize}
%The same reasoning to select vectors $\ve'_j$ is made for $({B'}_0,\oplus)$. The maximum number of iterations $\ell$ that can be performed equals the number of cosets, $2^{s}$. If this maximum is reached, designs $\mathcal{P}_\ell$ and $\mathcal{P}'_\ell$ both correspond to a full grid design of $2^{rs}$ points.
%We have the following result:
%\item[Step 4.] The two blocks $B_l$ and ${B'}_l$ are then both randomized using the same Owen's scrambling column-wise.
